Global search algorithms for one-dimensional multiextremal optimization problems with nonlinear constraints

### 2017-10-06 2017/ParOptBook/Chapter4.tex
parOptBook: eliminate all hardcoded references in chapter 4

Assume the constraint with the number $j,\: 1\le j\le m$, is fulfilled at all points of the domain
\begin{equation}
  Q_j=\{x\in[a,b]:g_j(x)\le 0\},\: 1\le j\le m,
\end{equation}
which hereinafter will be called the feasible domain for this constraint. The feasible domain $Q$ of initial problem \eqref{eq4:problem} can be defined by the condition
\begin{equation}
  \label{eq4:Q_union}
  Q=Q_1\cap\dots\cap Q_m.
\end{equation}

\textbf{The constraints’ check order.} he numbering of the constraints and the objective function introduced in the statement of problem \eqref{eq4:problem} and defined by the sequence of integer numbers
\begin{equation}
  \label{eq4:basic_h}
  H=\{1,2,\dots,m,m+1\},
\end{equation}
will be considered as a base one.
Assume that every particular point $x$ from the search domain $[a,b]$ can have its own order of the problem constraints’ checking. The order of such checking corresponding to the point $x_i$ from \eqref{eq4:characteristic} will be defines as some permutation $H(x_i )$ of the basic numbers from \eqref{eq4:basic_h}, i.e.,
\begin{equation}
  H(x_i)=\{j_{i1},\dots,j_{im},j_{i,m+1}\},\:0\le i\le k,
\end{equation}
where
\begin{equation}
  j_{i,m+1}=m+1,\:0\le i\le k.
\end{equation}

Since the constraints’ check order while the trial execution at the point $x$ i is determined by the corresponding tuple $H(x_i )$ from (4.33), the termination of the trial at this point after the computation of the values of l problem functions produces the value $z_i=g_\nu(x_i)$ where $\nu=j_{il}$. In this connection, let us use not the serial number $l$ of the last function calculated in this point, as it was accepted in \eqref{eq4:trial} but the index of the last function in the basic numeration (4.32) as the
value of $\nu(x_i )$ for the point $x_i$ , i.e.,
\begin{equation}
  \nu(x_i)=j_{il}.
\end{equation}

@@

\emph{Modification 1 (the fixation of the check order)}. Each point $x_i$ from \eqref{eq4:search_data} is assigned the corresponding checking order $H(x_i )$ from (4.33), and this order is included into the algorithm database.

\emph{Modification 2 (the determination of the trial points’ indices).} Each point $x_i$ from \eqref{eq4:search_data} is assigned the value $z_i=g_\nu(x_i)$ where according to (4.35), the index $\nu=\nu(x_i)$ is the basic index from (4.32) corresponding to the value of the last function estimated in this point. These indices are used in all computations based on expressions \eqref{fig:4_13}-\eqref{eq4:characteristic}, \eqref{eq4:next_point}.

The modifications in the method of using the indices are related to rules \eqref{eq4:characteristic} of calculations of the characteristics $R(i)$ of the intervals $(x_{i-1},x_i),\: 1\le i\le k$, with noncoincident
indices of the boundary points only. Assume
\begin{equation}
  \nu(x_{i-1})=j_{i-1,p}=j_{i,q}\not=\nu(x_i)=j_{i-1,u}=j_{i,x},
\end{equation}
i.e., the problem function with the basic index $\nu(x_{i-1})$ has the index $p$ in the sequence $H(x_{i-1})$ corresponding to the point $x_{i-1}$ and the index $q$ in the sequence $H(x_i)$ corresponding to the point $x_i$. Analogously, the function with the basic index $\nu(x_i)$ corresponds to the index $u$ from the sequence $H(x_{i-1})$ and the index s from the sequence $H(x_i)$. Then, the introduced modification altering the sixth rule of FCOM consists in the following.

If for the indices $p,q, u$ and $s$ from (4.36) the conditions
\begin{equation}
  p<u,q<s,
\end{equation}
are fulfilled, in order to determine the characteristic $R(i)$ one should use the second expression from formulae \eqref{eq4:characteristic} with $\nu=\nu(x_i)$ (note that in this case the value $g_\nu(x_{i-1})$ is undefined).

Otherwise, when the reverse to (4.37) conditions are fulfilled
\begin{equation}
  p>u,q>s,
\end{equation}
the characteristic $R(i)$ is defined by the third expression from \eqref{eq4:characteristic} with $\nu=\nu(x_{i-1})$ (in this
case, the value $g_\nu(x_i ) is undefined)$. In the cases when conditions (4.37) and (4.38) are not fulfilled, the initial order of calculations prescribed by the sixth rule of the initial algorithm remains.

\emph{Modification 3 (the choice of the constraints’ check order in the current trial point).} Assume that the constraints’ check order corresponding to the first two trial points is defined by
the conditions
\begin{equation}
  H(a)=H(b)=H,
\end{equation}
where $H$ from (4.32). Let
\begin{displaymath}
  p=\nu(x_{t-1}),\: q=\nu(x_t)
\end{displaymath}
are the indices of the boundary points of the interval $(x_{t-1},x_t)$ including the current trial point $x^{k+1}$ (according to the eighth rule) determined according to (4.35). Let us introduce a notation $\{s,H\backslash s\}$ for the tuple obtained from $H$ by excluding the index $s$, shifting all indices $1,2,\dots,s-1$ to the right of one position, and the transfer of the index s excluded from $H$ into the first position, i.e.,
\begin{displaymath}
  \{s,H\backslash s\}=\{s,1,2,\dots,s-1,s+1,\dots,q-1,q+1,\dots,m+1\}.
\end{displaymath}
When $1\le p\le q\le m$, the superposition of the operations of the type described above produces the tuple of the kind
\begin{displaymath}
  \{p,\{q,H\backslash q\}\backslash p\}=\{p,q,1,\dots,p-1,p+1,\dots,q-1,q+1,\dots,m+1\}.
\end{displaymath}
Note that according to (4.34), the index $m+1$ always occupies the last position.

Now let us agree to determine the order $H(x^{k+1}),\: k\ge 1$ according to the following rules:
\begin{gather}
 H(x^{k+1})=H,\: p=q=m+1, \nonumber \\
 H(x^{k+1})=\{p,H\backslash p\}, (p=q<m+1)\cup(p<q=m+1), \nonumber \\
 H(x^{k+1})=\{q,H\backslash q\}, p=m+1>q\\
 H(x^{k+1})=\{p, \{q,H\backslash q\}\backslash p\}, p<q\le m \nonumber \\
 H(x^{k+1})=\{q, \{p,H\backslash p\}\backslash q\} m\ge p>q. \nonumber
\end{gather}

\textbf{Note.} While solving the problem \eqref{eq4:problem} rules (4.39) and (4.40) included into the scheme of ACOM assign the index values from (4.35) to all trial points unambiguously. Then, any feasible point $x$ always corresponds to the index $\nu(x)=m+1$. This value can be predicted before the trial execution in such point as well. However, if the point $x$ is not a feasible one and does not belong to the points, in which the trials have been executed already, the index value can be undefined in this point.

@@

\begin{theorem}
Assume problem \eqref{eq4:problem} to have a solution $x^*$ and the following conditions are fulfilled:
\begin{enumerate}
  \item each domain $Q_j ,1\le j\le m+1$ from (4.30) represents a union of a finite number of segments with a positive length;
  \item each function $g_j(x): 1\le j\le m+1$ satisfies Lipschitz condition with the corresponding constant $L_j$;
  \item the components $\epsilon_\nu$ of the reserve vector $\epsilon_R$ satisfy the inequalities
  \begin{equation}
    0<2\epsilon_\nu<L_\nu(\beta-\alpha),
  \end{equation}
  where $\beta-\alpha$ is the length of the interval $[\alpha,\beta]$ belonging to the feasible domain $Q$ from \eqref{eq4:Q_union} and including the point $x^*$, which is the solution of the problem;
  \item since some sufficiently large step number $k$ the values $\mu_\nu$ from \eqref{eq4:lip_est} corresponding to the nonempty sets $I_\nu$ satisfy the inequalities
  \begin{displaymath}
    r_\nu\mu_\nu>2L_\nu,\: 1\le\nu\le m+1,
  \end{displaymath}
  where $r_\nu>1,\: 1\le\nu\le m+1$ are the parameters of the algorithm.
\end{enumerate}

@@

The distribution of the interval lengths $[\alpha,\beta]$ from (4.41) including the solution points $x^*$ corresponding to the problems from the first set is presented in Fig. \ref{fig:4_11}. Fig. \ref{fig:4_12} demonstrates a similar distribution for the problems of the second set. In these figures, the values $(\beta=\alpha)/(b-a)$ (in percentage) are plotted against the abscissa axes while the numbers of problems in the corresponding sets are plotted against the ordinate ones. In order to obtain the discussed estimates, the lengths of the intervals $[\alpha,\beta]$ corresponding to the respective functions were determined by a numerical method.

@@

\begin{lemma}
  (on the sequence of the shrinking intervals). Assume the conditions of Theorem \ref{th4:convergence_par} to be satisfied. Then for each limit point $x$ of the sequence $\{x_k\}$ there exists an infinite sequence of the intervals
  \begin{equation}
    \{(x_{t-1},x_t):t=t(q^s)\},\:s=1,2,\dots,
  \end{equation}
  satisfying the conditions
  \begin{gather}
    \overline x= \cap_{s=1}^\infty[x_{t-1},x_t], \\
    \lim_{p\to\infty}\Delta_t=0, \\
    R(t(q^s))>0,\:s=1,2,\dots, \\
    \lim_{p\to\infty}R(t(q^s))=0,
  \end{gather}
  where $q^1<q^2<\dots;\:R(t),\:\Delta_t$, and $t$ are from rules \eqref{eq4:characteristic_par} and \eqref{eq4:MaxR_par}, correspondingly.
\end{lemma}

@@

  It follows from this condition, from the rules of the next trial point selection \eqref{eq4:next_point_1_par} and \eqref{eq4:next_point_2_par}, and from the estimate of the value $\mu_\nu$ from \eqref{eq4:lip_est_par}, that
  \begin{equation}
    \label{eq4:4_60}
    \max\{x_t-x^{q+1},x^{q+1}-x_{t-1}\}\le\gamma \Delta_t,
  \end{equation}
  where $\gamma=(r+1)/2r<1$ and
  \begin{equation}
    r=\min\{r_\nu:1\le\nu\le m+1\}>1.
  \end{equation}
  \label{eq4:4_61}
  Then, the property (4.56) is a consequence of inequality \eqref{eq4:4_60}.

  At every step $k\ge 1$ there exists an interval $(x_{i-1},x_i), \:i=i(k)$, at one of the ends of which the value $z^*_\omega$ from \eqref{eq4:z_const_par} is achieved. This value corresponds to the index $\omega$ defined (when $\omega<m$) by the conditions $I_\omega\not= \varnothing,\: I_{\omega+1} =  \varnothing$. If $I_{m+1}= \varnothing$, the index $\omega=m+1$. If at that $\nu(x_{i -1} )\not = \nu(x_i )$, then according (4.47) and (4.50), $R(i)=2\Delta_i>0$. In the case when $\nu(x_{i -1} )= \nu(x_i )=\omega$, the estimate
  \begin{displaymath}
    \max(z_{i-1}, z_i)-z^*_\omega\le\mu\Delta_i
  \end{displaymath}
  follows from \eqref{eq4:lip_est_par} and, along with \eqref{eq4:characteristic_par}, gives the estimate
  \begin{displaymath}
    R(i)=\Delta_i\left(1-\frac{\max(z_{i-1},z_i)-z^*_\omega}{r_\omega\mu_\omega\Delta_i}\right)\ge (1-r^{-1})\Delta_i,
  \end{displaymath}
  where $r$ is from \eqref{eq4:4_61}. Thus, at any search step $k\ge 1$ there exists an interval having a strictly positive characteristic. Hence, taking into account the rules of selection of the interval with the greatest characteristic \eqref{eq4:MaxR_par}, one can conclude that the intervals from system (4.54) should also have the positive characteristics since these ones undergo splitting up by the trial points falling into these intervals. Consequently, conditions (4.57) hold.

  It follows from the trial execution rules and condition (4.35) that for $\nu(x_i )\le m$ the values $z_i$ assigned to the points $x_i$ from \eqref{eq4:par_search_data} are positive. Therefore, according to \eqref{eq4:z_const_par},
  \begin{displaymath}
    z_i=g_\nu(x_i)\ge z^*_nu,\:1\le\nu=\nu(x_i)\le m+1,\: 0\le i\le k.
  \end{displaymath}
    Now from the rules \eqref{eq4:characteristic_par} for the characteristics and the limit (4.56), taking into account the condition \eqref{eq4:lip_condition} of Lipschitzness of the functions and inequality (4.57), one can derive out the truth of statement (4.58).
\end{proof}
\begin{lemma}
  \label{lm4:2}
  (on the existence of a feasible point). Assume the conditions of Theorem \ref{th4:convergence_par} to be fulfilled. Then, there exists the trial index $h\ge 1$, for which
  \begin{equation}
    \label{eq4:62}
    \nu(x^h)=m+1.
  \end{equation}
\end{lemma}
\begin{proof}
  Let us assume the opposite and consider the interval $[x_{t-1},x_t]$ including the problem solution $x^*$ from \eqref{eq4:problem} at the step $k$, i.e., $t=t(k)$. Since we have assumed the equality \eqref{eq4:62} not to take place and, consequently, the sequence $\{x^k\}$ does not include the points from the feasible domain $Q$, the interval $[\alpha,\beta]$ from (4.52) should satisfy the conditions
  \begin{equation}
    \label{eq4:63}
    x^*\in [\alpha,\beta]\subset[x_{t-1}, x_t].
  \end{equation}
  Then, from this condition as well as from the Lipschitz properties of the problem functions the relations
  \begin{gather}
    \label{eq4:64}
    z_{t-1}\le L_\nu(\alpha - x_{t-1})\le L_\nu\Delta_t-L_\nu(\beta-\alpha),\:\nu=\nu(x_{t-1}) \\
    \label{eq4:65}
    z_t\le L_\nu(x_{t}-\beta)\le L_\nu\Delta_t-L_\nu(\beta-\alpha),\:\nu=\nu(x_{t})
  \end{gather}
  follow. It was taken into account when deriving out of the above relations that
  \begin{displaymath}
    g_\nu(x)\le 0,\: x\in[\alpha,\beta]\subset Q.
  \end{displaymath}
  Analogously, the case $\nu=\nu(x_{t-1})=\nu(x_t)$ corresponds to the inequality
  \begin{equation}
    \label{eq4:66}
    z_t+z_{t-1}\le L_\nu \Delta_t - L_\nu(\beta-\alpha).
  \end{equation}

@@

  From (4.58), \eqref{eq4:67}, and from the rule (4.48) determining the choice of the interval for the next trial point, the inevitability of carrying out a trial in the interval $[x_{t-1},x_t]$ from \eqref{eq4:63} follows. Consequently, the interval $[x_{t-1},x_t]$ will be divided into two subintervals. After that, at least one of those will include the point $x^*$. The repetition of this division generates (with $k\to\infty$) an infinite sequence of the subintervals shrinking to the point $x^*$ . In other words, if the third condition of Theorem \ref{th4:convergence_par} and the fourth one are satisfied, $x^*$ is the limit point of the sequence ${x^k}$.

  Since the coincidence $x^*$ with one of the trial points means \eqref{eq4:62} to hold, then it remains to consider the case when $x^*\in(x_{t- 1} ,x_t ),\: t=t(k)$ for any $k\ge 1$. Here, because of the first condition of Theorem \ref{th4:convergence_par} and the definition (4.31) either $[x_{t-1},x^*]\in Q$ or $[x^* ,x_t]\in Q$ if the number $k$ is sufficiently large. In this case, either $\nu(x_{t-1})=m+1$ or $\nu (x_t )=m+1.$
\end{proof}

@@

\begin{proof}
  Let us assume the opposite. According to (4.55) and (4.56),
  \begin{displaymath}
    \overline x = \lim_{s\to\infty}x_{t-1}=\lim_{s\to\infty}x_t, t=t(x^s).
  \end{displaymath}
  where q s is from (4.57) Therefore, it follows from the assumption $x\in Q$ that there exists such large number $S>1$ and such small subinterval
  \begin{equation}
    \label{eq4:69}
    (u,w)\subset[a,b]\setminus Q,
  \end{equation}
  that
  \begin{equation}
    \label{eq4:70}
    [x_{t-1},x_t]\subset(u,w),\:t=t(q^s),\:s\ge S.
  \end{equation}

  Let us assign to $Q_j, 1\le j\le m$, from (4.30) their open complements $D_j=[a,b]\setminus Q_j$ consisting of finite number of the open intervals with the positive length (as a consequence of the first condition of Theorem \ref{th4:convergence_par}). For these complements
  \begin{equation}
    \label{eq4:71}
    g_j(x)>0,x\in D_j,\:1\le j\le m.
  \end{equation}

@@

  According to the trial execution rules and to conditions (4.35) and \eqref{eq4:72}, for any interval from \eqref{eq4:70} the relations
  \begin{displaymath}
    \omega_1=\nu(x_{t-1})\in M,\: \omega_2=\nu(x_t)\in N,\:t=t(q^s),
  \end{displaymath}
  hold where $1\le \omega_1,\: \omega_2\le m$. Moreover, according to the rules (4.39) and (4.40) defining the constraints’ check order at trial points,
  \begin{displaymath}
    \nu(x_{t-1}),\:\nu(x_t)\in\{\omega_1,\omega_2\},\: t=t(q^{p+1}).
  \end{displaymath}
  In other words, for sufficiently large number $s>1$ the indices assigned to the boundary points of the considered system of intervals $[x_{t-1},x_t],\: t=t(q^s ): s\ge S$ can take the values from the set $\{\omega_1,\omega_2\}$ only.

>>>>

Assume the constraint with the number $j,\: 1\le j\le m$, is fulfilled at all points of the domain
\begin{equation}
  \label{eq4:30}
  Q_j=\{x\in[a,b]:g_j(x)\le 0\},\: 1\le j\le m,
\end{equation}
which hereinafter will be called the feasible domain for this constraint. The feasible domain $Q$ of initial problem \eqref{eq4:problem} can be defined by the condition
\begin{equation}
  \label{eq4:Q_union}
  Q=Q_1\cap\dots\cap Q_m.
\end{equation}

\textbf{The constraints’ check order.} The numbering of the constraints and the objective function introduced in the statement of problem \eqref{eq4:problem} and defined by the sequence of integer numbers
\begin{equation}
  \label{eq4:basic_h}
  H=\{1,2,\dots,m,m+1\},
\end{equation}
will be considered as a base one.
Assume that every particular point $x$ from the search domain $[a,b]$ can have its own order of the problem constraints’ checking. The order of such checking corresponding to the point $x_i$ from \eqref{eq4:characteristic} will be defines as some permutation $H(x_i )$ of the basic numbers from \eqref{eq4:basic_h}, i.e.,
\begin{equation}
  \label{eq4:33}
  H(x_i)=\{j_{i1},\dots,j_{im},j_{i,m+1}\},\:0\le i\le k,
\end{equation}
where
\begin{equation}
  \label{eq4:34}
  j_{i,m+1}=m+1,\:0\le i\le k.
\end{equation}

Since the constraints’ check order while the trial execution at the point $x$ i is determined by the corresponding tuple $H(x_i )$ from \eqref{eq4:33}, the termination of the trial at this point after the computation of the values of l problem functions produces the value $z_i=g_\nu(x_i)$ where $\nu=j_{il}$. In this connection, let us use not the serial number $l$ of the last function calculated in this point, as it was accepted in \eqref{eq4:trial} but the index of the last function in the basic numeration \eqref{eq4:basic_h} as the
value of $\nu(x_i )$ for the point $x_i$ , i.e.,
\begin{equation}
  \label{eq4:35}
  \nu(x_i)=j_{il}.
\end{equation}

@@

\emph{Modification 1 (the fixation of the check order)}. Each point $x_i$ from \eqref{eq4:search_data} is assigned the corresponding checking order $H(x_i )$ from \eqref{eq4:33}, and this order is included into the algorithm database.

\emph{Modification 2 (the determination of the trial points’ indices).} Each point $x_i$ from \eqref{eq4:search_data} is assigned the value $z_i=g_\nu(x_i)$ where according to \eqref{eq4:35}, the index $\nu=\nu(x_i)$ is the basic index from \eqref{eq4:basic_h} corresponding to the value of the last function estimated in this point. These indices are used in all computations based on expressions \eqref{fig:4_13}-\eqref{eq4:characteristic}, \eqref{eq4:next_point}.

The modifications in the method of using the indices are related to rules \eqref{eq4:characteristic} of calculations of the characteristics $R(i)$ of the intervals $(x_{i-1},x_i),\: 1\le i\le k$, with noncoincident
indices of the boundary points only. Assume
\begin{equation}
  \label{eq4:36}
  \nu(x_{i-1})=j_{i-1,p}=j_{i,q}\not=\nu(x_i)=j_{i-1,u}=j_{i,x},
\end{equation}
i.e., the problem function with the basic index $\nu(x_{i-1})$ has the index $p$ in the sequence $H(x_{i-1})$ corresponding to the point $x_{i-1}$ and the index $q$ in the sequence $H(x_i)$ corresponding to the point $x_i$. Analogously, the function with the basic index $\nu(x_i)$ corresponds to the index $u$ from the sequence $H(x_{i-1})$ and the index s from the sequence $H(x_i)$. Then, the introduced modification altering the sixth rule of FCOM consists in the following.

If for the indices $p,q, u$ and $s$ from \eqref{eq4:36} the conditions
\begin{equation}
  \label{eq4:37}
  p<u,q<s,
\end{equation}
are fulfilled, in order to determine the characteristic $R(i)$ one should use the second expression from formulae \eqref{eq4:characteristic} with $\nu=\nu(x_i)$ (note that in this case the value $g_\nu(x_{i-1})$ is undefined).

Otherwise, when the reverse to \eqref{eq4:37} conditions are fulfilled
\begin{equation}
  \label{eq4:38}
  p>u,q>s,
\end{equation}
the characteristic $R(i)$ is defined by the third expression from \eqref{eq4:characteristic} with $\nu=\nu(x_{i-1})$ (in this
case, the value $g_\nu(x_i ) is undefined)$. In the cases when conditions \eqref{eq4:37} and \eqref{eq4:38} are not fulfilled, the initial order of calculations prescribed by the sixth rule of the initial algorithm remains.

\emph{Modification 3 (the choice of the constraints’ check order in the current trial point).} Assume that the constraints’ check order corresponding to the first two trial points is defined by
the conditions
\begin{equation}
  \label{eq4:39}
  H(a)=H(b)=H,
\end{equation}
where $H$ from \eqref{eq4:basic_h}. Let
\begin{displaymath}
  p=\nu(x_{t-1}),\: q=\nu(x_t)
\end{displaymath}
are the indices of the boundary points of the interval $(x_{t-1},x_t)$ including the current trial point $x^{k+1}$ (according to the eighth rule) determined according to \eqref{eq4:35}. Let us introduce a notation $\{s,H\backslash s\}$ for the tuple obtained from $H$ by excluding the index $s$, shifting all indices $1,2,\dots,s-1$ to the right of one position, and the transfer of the index s excluded from $H$ into the first position, i.e.,
\begin{displaymath}
  \{s,H\backslash s\}=\{s,1,2,\dots,s-1,s+1,\dots,q-1,q+1,\dots,m+1\}.
\end{displaymath}
When $1\le p\le q\le m$, the superposition of the operations of the type described above produces the tuple of the kind
\begin{displaymath}
  \{p,\{q,H\backslash q\}\backslash p\}=\{p,q,1,\dots,p-1,p+1,\dots,q-1,q+1,\dots,m+1\}.
\end{displaymath}
Note that according to \eqref{eq4:34}, the index $m+1$ always occupies the last position.

Now let us agree to determine the order $H(x^{k+1}),\: k\ge 1$ according to the following rules:
\begin{gather}
 H(x^{k+1})=H,\: p=q=m+1, \nonumber \\
 H(x^{k+1})=\{p,H\backslash p\}, (p=q<m+1)\cup(p<q=m+1), \nonumber \\
 \label{eq4:40}
 H(x^{k+1})=\{q,H\backslash q\}, p=m+1>q\\
 H(x^{k+1})=\{p, \{q,H\backslash q\}\backslash p\}, p<q\le m \nonumber \\
 H(x^{k+1})=\{q, \{p,H\backslash p\}\backslash q\} m\ge p>q. \nonumber
\end{gather}

\textbf{Note.} While solving the problem \eqref{eq4:problem} rules \eqref{eq4:39} and \eqref{eq4:40} included into the scheme of ACOM assign the index values from \eqref{eq4:35} to all trial points unambiguously. Then, any feasible point $x$ always corresponds to the index $\nu(x)=m+1$. This value can be predicted before the trial execution in such point as well. However, if the point $x$ is not a feasible one and does not belong to the points, in which the trials have been executed already, the index value can be undefined in this point.

@@

\begin{theorem}
Assume problem \eqref{eq4:problem} to have a solution $x^*$ and the following conditions are fulfilled:
\begin{enumerate}
  \item each domain $Q_j ,1\le j\le m+1$ from \eqref{eq4:30} represents a union of a finite number of segments with a positive length;
  \item each function $g_j(x): 1\le j\le m+1$ satisfies Lipschitz condition with the corresponding constant $L_j$;
  \item the components $\epsilon_\nu$ of the reserve vector $\epsilon_R$ satisfy the inequalities
  \begin{equation}
    \label{eq4:41}
    0<2\epsilon_\nu<L_\nu(\beta-\alpha),
  \end{equation}
  where $\beta-\alpha$ is the length of the interval $[\alpha,\beta]$ belonging to the feasible domain $Q$ from \eqref{eq4:Q_union} and including the point $x^*$, which is the solution of the problem;
  \item since some sufficiently large step number $k$ the values $\mu_\nu$ from \eqref{eq4:lip_est} corresponding to the nonempty sets $I_\nu$ satisfy the inequalities
  \begin{displaymath}
    r_\nu\mu_\nu>2L_\nu,\: 1\le\nu\le m+1,
  \end{displaymath}
  where $r_\nu>1,\: 1\le\nu\le m+1$ are the parameters of the algorithm.
\end{enumerate}

@@

The distribution of the interval lengths $[\alpha,\beta]$ from \eqref{eq4:41} including the solution points $x^*$ corresponding to the problems from the first set is presented in Fig. \ref{fig:4_11}. Fig. \ref{fig:4_12} demonstrates a similar distribution for the problems of the second set. In these figures, the values $(\beta=\alpha)/(b-a)$ (in percentage) are plotted against the abscissa axes while the numbers of problems in the corresponding sets are plotted against the ordinate ones. In order to obtain the discussed estimates, the lengths of the intervals $[\alpha,\beta]$ corresponding to the respective functions were determined by a numerical method.

@@

\begin{lemma}
  (on the sequence of the shrinking intervals). Assume the conditions of Theorem \ref{th4:convergence_par} to be satisfied. Then for each limit point $x$ of the sequence $\{x_k\}$ there exists an infinite sequence of the intervals
  \begin{equation}
    \label{eq4:54}
    \{(x_{t-1},x_t):t=t(q^s)\},\:s=1,2,\dots,
  \end{equation}
  satisfying the conditions
  \begin{gather}
    \label{eq4:55}
    \overline x= \cap_{s=1}^\infty[x_{t-1},x_t], \\
    \label{eq4:56}
    \lim_{p\to\infty}\Delta_t=0, \\
    \label{eq4:57}
    R(t(q^s))>0,\:s=1,2,\dots, \\
    \label{eq4:58}
    \lim_{p\to\infty}R(t(q^s))=0,
  \end{gather}
  where $q^1<q^2<\dots;\:R(t),\:\Delta_t$, and $t$ are from rules \eqref{eq4:characteristic_par} and \eqref{eq4:MaxR_par}, correspondingly.
\end{lemma}

@@

  It follows from this condition, from the rules of the next trial point selection \eqref{eq4:next_point_1_par} and \eqref{eq4:next_point_2_par}, and from the estimate of the value $\mu_\nu$ from \eqref{eq4:lip_est_par}, that
  \begin{equation}
    \label{eq4:4_60}
    \max\{x_t-x^{q+1},x^{q+1}-x_{t-1}\}\le\gamma \Delta_t,
  \end{equation}
  where $\gamma=(r+1)/2r<1$ and
  \begin{equation}
    r=\min\{r_\nu:1\le\nu\le m+1\}>1.
  \end{equation}
  \label{eq4:4_61}
  Then, the property \eqref{eq4:56} is a consequence of inequality \eqref{eq4:4_60}.

  At every step $k\ge 1$ there exists an interval $(x_{i-1},x_i), \:i=i(k)$, at one of the ends of which the value $z^*_\omega$ from \eqref{eq4:z_const_par} is achieved. This value corresponds to the index $\omega$ defined (when $\omega<m$) by the conditions $I_\omega\not= \varnothing,\: I_{\omega+1} =  \varnothing$. If $I_{m+1}= \varnothing$, the index $\omega=m+1$. If at that $\nu(x_{i -1} )\not = \nu(x_i )$, then according \eqref{eq4:characteristic_par} and \eqref{eq4:next_point_2_par}, $R(i)=2\Delta_i>0$. In the case when $\nu(x_{i -1} )= \nu(x_i )=\omega$, the estimate
  \begin{displaymath}
    \max(z_{i-1}, z_i)-z^*_\omega\le\mu\Delta_i
  \end{displaymath}
  follows from \eqref{eq4:lip_est_par} and, along with \eqref{eq4:characteristic_par}, gives the estimate
  \begin{displaymath}
    R(i)=\Delta_i\left(1-\frac{\max(z_{i-1},z_i)-z^*_\omega}{r_\omega\mu_\omega\Delta_i}\right)\ge (1-r^{-1})\Delta_i,
  \end{displaymath}
  where $r$ is from \eqref{eq4:4_61}. Thus, at any search step $k\ge 1$ there exists an interval having a strictly positive characteristic. Hence, taking into account the rules of selection of the interval with the greatest characteristic \eqref{eq4:MaxR_par}, one can conclude that the intervals from system \eqref{eq4:54} should also have the positive characteristics since these ones undergo splitting up by the trial points falling into these intervals. Consequently, conditions \eqref{eq4:57} hold.

  It follows from the trial execution rules and condition \eqref{eq4:35} that for $\nu(x_i )\le m$ the values $z_i$ assigned to the points $x_i$ from \eqref{eq4:par_search_data} are positive. Therefore, according to \eqref{eq4:z_const_par},
  \begin{displaymath}
    z_i=g_\nu(x_i)\ge z^*_nu,\:1\le\nu=\nu(x_i)\le m+1,\: 0\le i\le k.
  \end{displaymath}
    Now from the rules \eqref{eq4:characteristic_par} for the characteristics and the limit \eqref{eq4:56}, taking into account the condition \eqref{eq4:lip_condition} of Lipschitzness of the functions and inequality \eqref{eq4:57}, one can derive out the truth of statement \eqref{eq4:58}.
\end{proof}
\begin{lemma}
  \label{lm4:2}
  (on the existence of a feasible point). Assume the conditions of Theorem \ref{th4:convergence_par} to be fulfilled. Then, there exists the trial index $h\ge 1$, for which
  \begin{equation}
    \label{eq4:62}
    \nu(x^h)=m+1.
  \end{equation}
\end{lemma}
\begin{proof}
  Let us assume the opposite and consider the interval $[x_{t-1},x_t]$ including the problem solution $x^*$ from \eqref{eq4:problem} at the step $k$, i.e., $t=t(k)$. Since we have assumed the equality \eqref{eq4:62} not to take place and, consequently, the sequence $\{x^k\}$ does not include the points from the feasible domain $Q$, the interval $[\alpha,\beta]$ from \eqref{eq4:reserves_par} should satisfy the conditions
  \begin{equation}
    \label{eq4:63}
    x^*\in [\alpha,\beta]\subset[x_{t-1}, x_t].
  \end{equation}
  Then, from this condition as well as from the Lipschitz properties of the problem functions the relations
  \begin{gather}
    \label{eq4:64}
    z_{t-1}\le L_\nu(\alpha - x_{t-1})\le L_\nu\Delta_t-L_\nu(\beta-\alpha),\:\nu=\nu(x_{t-1}) \\
    \label{eq4:65}
    z_t\le L_\nu(x_{t}-\beta)\le L_\nu\Delta_t-L_\nu(\beta-\alpha),\:\nu=\nu(x_{t})
  \end{gather}
  follow. It was taken into account when deriving out of the above relations that
  \begin{displaymath}
    g_\nu(x)\le 0,\: x\in[\alpha,\beta]\subset Q.
  \end{displaymath}
  Analogously, the case $\nu=\nu(x_{t-1})=\nu(x_t)$ corresponds to the inequality
  \begin{equation}
    \label{eq4:66}
    z_t+z_{t-1}\le L_\nu \Delta_t - L_\nu(\beta-\alpha).
  \end{equation}

@@

  From \eqref{eq4:58}, \eqref{eq4:67}, and from the rule \eqref{eq4:MaxR_par} determining the choice of the interval for the next trial point, the inevitability of carrying out a trial in the interval $[x_{t-1},x_t]$ from \eqref{eq4:63} follows. Consequently, the interval $[x_{t-1},x_t]$ will be divided into two subintervals. After that, at least one of those will include the point $x^*$. The repetition of this division generates (with $k\to\infty$) an infinite sequence of the subintervals shrinking to the point $x^*$ . In other words, if the third condition of Theorem \ref{th4:convergence_par} and the fourth one are satisfied, $x^*$ is the limit point of the sequence ${x^k}$.

  Since the coincidence $x^*$ with one of the trial points means \eqref{eq4:62} to hold, then it remains to consider the case when $x^*\in(x_{t- 1} ,x_t ),\: t=t(k)$ for any $k\ge 1$. Here, because of the first condition of Theorem \ref{th4:convergence_par} and the definition \eqref{eq4:Q_union} either $[x_{t-1},x^*]\in Q$ or $[x^* ,x_t]\in Q$ if the number $k$ is sufficiently large. In this case, either $\nu(x_{t-1})=m+1$ or $\nu (x_t )=m+1.$
\end{proof}

@@

\begin{proof}
  Let us assume the opposite. According to \eqref{eq4:55} and \eqref{eq4:56},
  \begin{displaymath}
    \overline x = \lim_{s\to\infty}x_{t-1}=\lim_{s\to\infty}x_t, t=t(x^s).
  \end{displaymath}
  where q s is from \eqref{eq4:57} Therefore, it follows from the assumption $x\in Q$ that there exists such large number $S>1$ and such small subinterval
  \begin{equation}
    \label{eq4:69}
    (u,w)\subset[a,b]\setminus Q,
  \end{equation}
  that
  \begin{equation}
    \label{eq4:70}
    [x_{t-1},x_t]\subset(u,w),\:t=t(q^s),\:s\ge S.
  \end{equation}

  Let us assign to $Q_j, 1\le j\le m$, from \eqref{eq4:30} their open complements $D_j=[a,b]\setminus Q_j$ consisting of finite number of the open intervals with the positive length (as a consequence of the first condition of Theorem \ref{th4:convergence_par}). For these complements
  \begin{equation}
    \label{eq4:71}
    g_j(x)>0,x\in D_j,\:1\le j\le m.
  \end{equation}

@@

  According to the trial execution rules and to conditions \eqref{eq4:35} and \eqref{eq4:72}, for any interval from \eqref{eq4:70} the relations
  \begin{displaymath}
    \omega_1=\nu(x_{t-1})\in M,\: \omega_2=\nu(x_t)\in N,\:t=t(q^s),
  \end{displaymath}
  hold where $1\le \omega_1,\: \omega_2\le m$. Moreover, according to the rules \eqref{eq4:39} and \eqref{eq4:40} defining the constraints’ check order at trial points,
  \begin{displaymath}
    \nu(x_{t-1}),\:\nu(x_t)\in\{\omega_1,\omega_2\},\: t=t(q^{p+1}).
  \end{displaymath}
  In other words, for sufficiently large number $s>1$ the indices assigned to the boundary points of the considered system of intervals $[x_{t-1},x_t],\: t=t(q^s ): s\ge S$ can take the values from the set $\{\omega_1,\omega_2\}$ only.
